% Paper cover page
\papertitle{Lessons from a blind study of simulated lenses: \\image reconstructions do not always reproduce true convergence}
\capauthors{
    \chapterauthor[1,2]{Philipp Denzel}
    \chapterauthor[3]{Sampath Mukherjee}
    \chapterauthor[4]{Jonathan P. Coles}
    \chapterauthor[2,1]{Prasenjit Saha}
}
\affils{
    \chapteraffil[1]{Institute for Computational Science, University of Zurich, 8057 Zurich, Switzerland}
    \chapteraffil[2]{Physics Institute, University of Zurich, 8057 Zurich, Switzerland}
    \chapteraffil[3]{STAR Institute, Quartier Agora - All\'ee du six Ao$\hat{u}$t, 19c B-4000 Li\`ege, Belgium}
    \chapteraffil[4]{Physik-Department, Technische Universit\"at M\"unchen, James-Franck-Str.~1, 85748 Garching, Germany}
}

\publishedin{Denzel20}
\clearpage


\def\pwidth{.66\textwidth}
\def\qwidth{.49\textwidth}
\def\pheight{.19\textheight}


% Abstract
\section*{Abstract}
  %
  \noindent In the coming years, strong gravitational lens discoveries are
  expected to increase in frequency by two orders of magnitude.  Lens-modelling
  techniques are being developed to prepare for the coming massive influx of new
  lens data, and blind tests of lens reconstruction with simulated data are
  needed for validation.  In this paper we present a systematic blind study of a
  sample of 15 simulated strong gravitational lenses from the EAGLE suite of
  hydrodynamic simulations.  We model these lenses with a free-form technique
  and evaluate reconstructed mass distributions using criteria based on shape,
  orientation, and lensed image reconstruction.  Especially useful is a lensing
  analogue of the Roche potential in binary star systems, which we call the {\em
  lensing Roche potential}.  This we introduce in order to factor out the
  well-known problem of steepness or mass-sheet degeneracy.  Einstein radii are
  on average well recovered with a relative error of ${\sim}5\%$ for quads and
  ${\sim}25\%$ for doubles; the position angle of ellipticity is on average also
  reproduced well up to $\pm10\degree$, but the reconstructed mass maps tend to
  be too round and too shallow.  It is also easy to reproduce the lensed images,
  but optimizing on this criterion does not guarantee better reconstruction of
  the mass distribution.


\clearpage
% Sections

% Introduction
\input{\home/tex/intro}

% Theory
\input{\home/tex/theory}

% % Methods
\input{\home/tex/methods}

% Results
\input{\home/tex/results}

% Conclusion
\input{\home/tex/conclusion}

%%%% Funding statement
% 
\section*{Funding statement}
PD acknowledges support from the Swiss National Science Foundation.
SM acknowledges the funding from the European Research Council (ERC)
under the EU’s Horizon 2020 research and innovation programme
(COSMICLENS;\@ grant agreement no.~787886).

%%%% Acknowledgements
% 
% \section*{Acknowledgments}
%

% Figures
\input{\home/fig/figures}
