% Paper cover page
\papertitle{The lens SW05 J143454.4+522850: a fossil group at redshift 0.6?}
\capauthors{
    \chapterauthor[1,2]{Philipp Denzel}
    \chapterauthor[2,1]{Prasenjit Saha}
}
\collabs{
    \chapterauthor[3]{Xanthe Gwyn Palmer}
    \chapterauthor[1]{Onur \c{C}atmabacak}
    \chapterauthor[4]{Jonathan Coles}
    \chapterauthor[5]{Claude Cornen}
    \chapterauthor[6]{Ignacio Ferreras}
    \chapterauthor[1]{Robert Feldmann}
    \chapterauthor[2]{Rafael K\"ueng}
    \chapterauthor[7]{Dominik Leier}
    \chapterauthor[3]{Aprajita Verma}


}
\affils{
    \chapteraffil[1]{Institute for Computational Science, University of Zurich, 8057 Zurich, Switzerland}
    \chapteraffil[2]{Physics Institute, University of Zurich, 8057 Zurich, Switzerland}
    \chapteraffil[3]{Sub-department of Astrophysics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford, OX1 3RH, UK}
    \chapteraffil[4]{Physik-Department, Technische Universit\"at M\"unchen, James-Franck-Str.~1, 85748 Garching, Germany}
    \chapteraffil[5]{Zooniverse, c/o Astrophysics Department, University of Oxford, Oxford OX1~3RH, UK}
    \chapteraffil[6]{Mullard Space Science Laboratory, University College London, Holmbury St Mary, Dorking, Surrey RH5 6NT, UK}
    \chapteraffil[7]{DB, Germany}
}

\publishedin[To be submitted to:\\\textit{Monthly Notices of the Royal Astronomical Society}]{}
\clearpage

\newcommand{\SW}{Space\,Warps\xspace}
\def\pwidth{.99\textwidth}
\def\qwidth{.49\textwidth}
\def\qqwidth{.57\textwidth}

% Abstract
\section*{Abstract}
\noindent Fossil groups are considered to be the end product of natural galaxy
group evolution in which group members sink towards the center of the
gravitational potential due to dynamical friction, and finally merge into a
single, massive, and (x-ray) bright elliptical.  Since gravitational lensing
depends on the mass of a foreground object, its mass concentration, and distance
to the observer, we can expect lensing effects of such fossil groups to be
particularly strong.

\noindent In this paper, we consider the exceptional system J143454.4$+$522850,
labelled SW05 by the \SW~discovery service.  We combine lens models with
population-synthesis models for the stellar mass to separate the total mass of
the lens into the two main components in galaxies, stars and dark matter,
thereby producing dark matter-maps for the lens.  The observed enclosed mass
profiles were additionally compared with state-of-the-art galaxy formation
simulations, which shows interesting differences between observation and
simulations.  These differences, as well as the lens model on itself, finally
lead us to the conclusion that SW05 may be a fossil group.  It would represent
the first direct comparison of a fossil groups' dark matter halo with its
stellar component in the center.  We finally put forth the idea of a diagnostic
for the identification of fossil groups through gravitational lensing by looking
for the lack of certain features in the lens models.
%

\clearpage
% Sections
  
% Introduction
\input{\home/tex/intro}
  
% System
\input{\home/tex/system}
  
% Lens mass model
\input{\home/tex/lensmass}
  
% Stellar mass model
\input{\home/tex/stelmass}
  
% Comparison to simulations
\input{\home/tex/simul}

% Conclusion
\input{\home/tex/disqus}

%%%% Acknowledgements
% 
\section*{Acknowledgments}
PhD acknowledges support from the Swiss National Science Foundation.