% Paper cover page
\papertitle{A new strategy for matching observed and simulated lensing galaxies}
\capauthors{
    \chapterauthor[1,2]{Philipp Denzel}
    \chapterauthor[3]{Sampath Mukherjee}
    \chapterauthor[2,1]{Prasenjit Saha}
}
\affils{
    \chapteraffil[1]{Institute for Computational Science, University of Zurich, 8057 Zurich, Switzerland}
    \chapteraffil[2]{Physics Institute, University of Zurich, 8057 Zurich, Switzerland}
    \chapteraffil[3]{STAR Institute, Quartier Agora - All\'ee du six Ao$\hat{u}$t, 19c B-4000 Li\`ege, Belgium}
}

%\publishedin[Reference:\\]{}
\clearpage


\newcommand\SDSSJ[1]{\ifthenelse{\equal{#1}{0029}}{\textsc{J0029\textminus0055}}{}\ifthenelse{\equal{#1}{0737}}{\textsc{J0737+3216}}{}\ifthenelse{\equal{#1}{0753}}{\textsc{J0753+3416}}{}\ifthenelse{\equal{#1}{1051}}{\textsc{J1051+4439}}{}\ifthenelse{\equal{#1}{0956}}{\textsc{J0956+5100}}{}\ifthenelse{\equal{#1}{1430}}{\textsc{J1430+6104}}{}\ifthenelse{\equal{#1}{1627}}{\textsc{J1627\textminus0053}}{}\ignorespaces}%
\newcommand\roche{\mathcal{P}}
\renewcommand{\tref}[1]{\hyperlink{tr#1}{(#1)}}
\renewcommand{\tlink}[1]{\hypertarget{tr#1}{(#1)}}



\def\pwidth{.99\textwidth}
\def\qheight{.23\textheight}


% Abstract
\section*{Abstract}
\noindent The study of strong-lensing systems conventionally involves
constructing a mass distribution that can reproduce the observed
multiply-imaging properties.  Such mass reconstructions are
generically non-unique.  Here we present an alternative strategy:
instead of modelling the mass distribution, we search cosmological
galaxy-formation simulations for plausible matches.  In this paper we
test the idea on seven well-studied lenses from the SLACS survey.  For
each of these, we first pre-select a few hundred galaxies from the
EAGLE simulations, using the expected Einstein radius as an initial
criterion.  Then, for each of these pre-selected galaxies, we fit for
the source light distribution, while using MCMC for the placement and
orientation of the lensing galaxy, so as to reproduce the multiple
images and arcs.  The results indicate that the strategy is feasible,
and even yields relative posterior probabilities of two different
galaxy-formation scenarios, though these are not statistically
significant yet.  Extension to other observables, such as kinematics
and colours of the stellar population in the lensing galaxy, is
straightforward in principle, though we have not attempted it yet.
Scaling to arbitrarily large numbers of lenses also appears feasible.
This will be especially relevant for upcoming wide-field surveys,
through which the number of galaxy lenses will rise possibly a
hundredfold, which will overwhelm conventional modelling methods.

\clearpage


% Introduction
\input{\home/tex/intro.tex}
  
% Methods
\input{\home/tex/methods.tex}

% SEAGLE
\input{\home/tex/seagle.tex}

% Test cases
\input{\home/tex/testcases.tex}

% Results
\input{\home/tex/results.tex}

% Conclusion
\input{\home/tex/conclusion.tex}


%%%% Acknowledgements
\section*{Acknowledgments}

  We would like to thank Liliya L. R. Williams and Dominique Sluse for useful
  discussions and suggestions on how to improve this paper.

  % We also thank the anonymous referee for the constructive suggestions to
  % bring the paper to its final form.

  PD acknowledges support from the Swiss National Science Foundation.  SM
  acknowledges the funding from the European Research Council (ERC) under the
  EUs Horizon 2020 research and innovation programme (COSMICLENS; grant
  agreement no. 787886).

  This research is based on observations made with the NASA/ESA Hubble Space
  Telescope obtained from the Space Telescope Science Institute, which is
  operated by the Association of Universities for Research in Astronomy, Inc.,
  under NASA contract NAS 5–26555. These observations are associated with
  programs \#10886, \#10174, \#12210, \#10494.

\section*{Data availability}
  The data underlying this article are available at the STScI
  (\href{https://mast.stsci.edu/}{https://mast.stsci.edu/}; the unique
  identifiers are cited in the acknowledgements).  The derived data generated in
  this research will be shared on request to the corresponding author, or can be
  replicated using the open-source software available at:
  \faGithub\;\href{https://github.com/phdenzel/gleam}{https://github.com/phdenzel/gleam}.


% Figures
\input{\home/fig/figures}