% Paper cover page
\papertitle{The Hubble constant from eight time-delay galaxy lenses}
\capauthors{
    \chapterauthor[1,2]{Philipp Denzel}
    \chapterauthor[3]{Jonathan P. Coles}
    \chapterauthor[4]{Liliya L. R. Williams}
    \chapterauthor[2,1]{Prasenjit Saha}
}
\affils{
    \chapteraffil[1]{Institute for Computational Science, University of Zurich, 8057 Zurich, Switzerland}
    \chapteraffil[2]{Physics Institute, University of Zurich, 8057 Zurich, Switzerland}
    \chapteraffil[3]{Physik-Department, Technische Universit\"at M\"unchen, James-Franck-Str.~1, 85748 Garching, Germany}
    \chapteraffil[4]{School of Physics and Astronomy, University of Minnesota, 116 Church Street SE, Minneapolis, MN 55455, USA}
}

\publishedin[Reference:\\]{Denzel20b}
\clearpage


\newcommand{\Hres}{\ensuremath{71.8^{+3.9}_{-3.3}\,\Hunits{}}}
\newcommand{\HaHzres}{\ensuremath{2.33^{+0.13}_{-0.11}\,\mathrm{aHz}}}
\newcommand{\invHres}{\ensuremath{13.7^{+0.7}_{-0.7}\,\Gyrs{}}}
\newcommand{\rhocritres}{\ensuremath{5.4^{+0.6}_{-0.5}\,\GeVmcube{}}}
\newcommand{\tref}[1]{\hyperlink{tr#1}{(#1)}}
\newcommand{\tlink}[1]{\hypertarget{tr#1}{(#1)}}
\def\twidth{.59\textwidth}
\def\theight{.59\textheight}
\def\pwidth{.49\textwidth}
\def\pheight{.49\textheight}
\def\qwidth{.49\textwidth}
\def\qheight{.24\textheight}


% Abstract
\section*{Abstract}
  %
  \noindent 
  We present a determination of the Hubble constant from the joint, free-form
  analysis of 8 strongly, quadruply lensing systems.  
  In the concordance cosmology, we find $\Ho{} = \Hres{}$ with a precision of
  $4.97\%$.  This is in agreement with the latest measurements from Supernovae
  Type Ia and Planck observations of the cosmic microwave background.  
  Our precision is lower compared to these and other recent time-delay
  cosmography determinations, because our modelling strategies reflect the
  systematic uncertainties of lensing degeneracies.  We furthermore are able to
  find reasonable lensed image reconstructions by constraining to either value
  of $\Ho$ from local and early Universe measurements.  This leads us to
  conclude that current lensing constraints on $\Ho$ are not strong enough to
  break the ``Hubble tension'' problem of cosmology.


  \clearpage
  % Sections
  
  % Introduction
  \input{\home/tex/intro}
  
  % Systems
  \input{\home/tex/systems}
  
  % Methods
  \input{\home/tex/methods}
  
  % Results
  \input{\home/tex/results}
  
  % Conclusion
  \input{\home/tex/conclusion}

  %%%% Acknowledgements
  % 
  \section*{Acknowledgments}
  PD acknowledges support from the Swiss National Science Foundation.  This
  research is based on observations made with the NASA/ESA Hubble Space
  Telescope obtained from the Space Telescope Science Institute, which is
  operated by the Association of Universities for Research in Astronomy, Inc.,
  under NASA contract NAS 5–26555. These observations are associated with
  programs \#10158, \#15320, \#12324, \#12874, \#9744, and \#10509.

  \section*{Data availability}
  The data underlying this article are available at the STScI
  (\href{https://mast.stsci.edu/}{https://mast.stsci.edu/}; the unique
  identifiers are cited in the acknowledgements).  The derived data generated in
  this research will be shared on request to the corresponding author.
  
  %
  % Figures
  \input{\home/fig/figures}
